---
title: "Activity 2: Simple Linear Regression"
author: "Sayed Muhammad Saifuddin"
date: "1/24/2023"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Day 1

```{r Packages, Task 2}
library(tidyverse)
library(tidymodels)
```

```{r Task3 Data Loading}

hfi <- readr::read_csv("https://www.openintro.org/data/csv/hfi.csv")
```

```{r Task3 Data Filtering}
hfi_2016 <- hfi %>% filter(year==2016)
hfi_2016 %>% select(pf_score, pf_expression_control)

```

## Simple Linear Regression

Below is a scatterplot of the two variables (note that the independent is on the horizontal axis and the dependent is on the vertical axis).

```{r Task3 Plotting relationship}

library(ggplot2)

# Basic scatter plot with labels
scatplt <- ggplot(hfi_2016, aes(x=pf_expression_control, y=pf_score)) + geom_point(size=2, color="blue")
scatplt <- scatplt + ggtitle("The relationship between personal freedom score & expression control \n ") +
  xlab("Personal Freedom Expression Control") + ylab("Personal Freedom Score")

# Change the color, size and face of the title, x and y axis labels

scatplt <- scatplt + theme(
plot.title = element_text(color="red", size=14, 
                          face="bold.italic", hjust = 0.5),
axis.title.x = element_text(color="black", size=14, face="bold"),
axis.title.y = element_text(color="black", size=14, face="bold")
)

scatplt

```

This data appears to follow a linear trend, and some R output for placing a linear model on this scatterplot, using regression analysis, is given below:

```{r alternative to Task4}
scatplt <- scatplt + geom_smooth(method=lm, se=FALSE, color="black")

scatplt

```

At this point, we want to build the actual model and examine characteristics that can be used to access the model. This can be accomplished in R using the linear model command "lm".

```{r Task5}
m1 <- lm(pf_score~pf_expression_control, data=hfi_2016)

#summary(m1)
tidy(m1)

```

# Day 2

## Correlation Coefficients

```{r Task2}
cor(hfi_2016[,c("pf_expression_control","pf_score")], use="everything")
```

```{r Task2}
glance(m1)
```

```{r Task3 Plotting relationship}

library(ggplot2)

# Basic scatter plot with labels
scatplt2 <- ggplot(hfi_2016, aes(x=pf_expression_control, y=pf_score)) +
  geom_smooth(method="lm", se = FALSE) + 
  geom_point(size=2, color="black")
scatplt2 <- scatplt2 + ggtitle("The relationship between personal freedom score & expression control \n ") +
  xlab("Personal Freedom Expression Control") + ylab("Personal Freedom Score")

# Change the color, size and face of the title, x and y axis labels

scatplt2 <- scatplt2 + theme(
plot.title = element_text(color="red", size=14,
                          face="bold.italic", hjust = 0.5),
axis.title.x = element_text(color="black", size=14, face="bold"),
axis.title.y = element_text(color="black", size=14, face="bold")
)

scatplt2

```

## Fitted (predicted) values and the residuals

```{r Task4 Model diagnostics}
m1_aug <- augment(m1)
m1_aug
```

## Plot of the residuals vs. fitted (predicted) values

```{r Task4 Model diagnostics}
ggplot(data = m1_aug, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted values") +
  ylab("Residuals")
```


## Histogram of the residuals

```{r Task4 Model diagnostics}
ggplot(data = m1_aug, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")
```

## Residual Analysis

```{r }

library(lmtest, pos=4)

bptest(pf_score ~ pf_expression_control, varformula = ~ fitted.values(m1), studentize=TRUE, data=hfi_2016)

plot(m1)

```

## Confidence Interval for the Slope: 

```{r}

confint(m1, level=0.95)

```

